---
title: "Using Monte Carlo Method to calculate Pi"
format:
  pdf:
    theme: cerulean
    number-sections: true
editor: 
  markdown: 
    wrap: 72
---
# Theory
We will calculate an approximation to $\pi$ using the Monte Carlo method.

We know the formula for the area of a circle: $$A_{circle} = \pi r^2$$
Also, we know the area formula for a square with side length  $d$:$$A_{square} = d^2$$ 
If we choose $d$ so that $2r=d$ we get the ratio of the areas:$$\frac{A_{circle}}{A_{square}} = \frac{\pi r^2}{4r^2} = \frac{\pi}{4}$$
Since we don't know $\pi$ we cannot calculate $A_{circle}$. But we can take advantage of random numbers and the implicit formula of the circle 
$x^2+y^2=1$ in a Cartesian coordinate system. 

We draw pairs of random numbers from $0$ to $1$ representing its position on the $x$- and the $y$-axis.Then we evaluate whether the $(x,y)$-pair is in the inner circle. For a large enough number of pairs, we get useful approximation to the above ratio. If we multiply it with $4$, we get an approximation of $\pi$.

\newpage
# Calculation in R
```{r calculation}
set.seed(3141)
n <- 1000
x <- runif(n)
y <- runif(n)
inner <- x^2 + y^2 <= 1
MC_pi <- 4*sum(inner)/length(inner)

MC_pi
```

\newpage
We will plot only a sample:

```{r circle, fig.height=5.558, fig.width=5}
library("plotrix")
plot(x, y, xlim=c(-0.1, 1.1), ylim=c(-0.1, 1.1), pch = "°")

draw.circle(0.5, 0.5, 0.5, border="red")

lines(x=c(0,1), y=c(0,0), col="red")
lines(x=c(0,1), y=c(1,1), col="red")
lines(x=c(0,0), y=c(0,1), col="red")
lines(x=c(1,1), y=c(1,0), col="red")
```


\newpage
# How many pairs do we need?
```{r pairs, fig.height=5.558, fig.width=5}
steps <- 1:1000
MC_pi.2 <- lapply(steps, function(x)
  4*sum(runif(x)^2+runif(x)^2 <= 1)/x)

plot(steps, MC_pi.2, type="l", ylab="approximation for Pi", 
     xlab="number of simulated (x,y)-pairs")
abline(h=pi, col="red")
```


